Ellipsoids (v1.0): 3D Magnetic modelling of ellipsoidal bodies

### 2017-04-20 manuscript/manuscript.tex

Let $(x, y, z)$ be a point referred to a Cartesian coordinate system with axes
$x$, $y$ and $z$ pointing to, respectively, North, East and down.
For convenience, we denominate this coordinate system as
\textit{main coordinate system} (Fig. \ref{fig:orientation-angles}a).
Let us consider an ellipsoidal body with centre at the point
$(x_{c}, y_{c}, z_{c})$, orientation defined by three angles
$\varepsilon$, $\zeta$, and $\eta$ (Fig. \ref{fig:orientation-angles}a),
and semi-axes defined by
positive constants $a$, $b$, $c$ (Fig. \ref{fig:orientation-angles}b).
The angles $\varepsilon$, $\zeta$, and $\eta$ are called \textit{strike},
\textit{dip}, and \textit{rake}, respectively, and are commonly used to
define the orientation of lines in structural geology
\citep{clark1986, allmendinger2012}.
The points $(x, y, z)$ located on the surface of this ellipsoidal body
satisfy the following equation:
\begin{equation}
(\mathbf{r} - \mathbf{r}_c)^T \mathbf{A} (\mathbf{r} - \mathbf{r}_c) = 1 \: ,
\label{eq:ellipsoid_surface}
\end{equation}
where $\mathbf{r} = [\begin{array}{ccc} x & y & z \end{array} ]^{\top}$,
$\mathbf{r}_{c} = [\begin{array}{ccc} x_{c} & y_{c} & z_{c} \end{array} ]^{\top}$,
$\mathbf{A}$ is a positive definite matrix given by
\begin{equation}
\mathbf{A} = \mathbf{V}
\left[ \begin{array}{ccc}
a^{-2} & 0 & 0 \\
0 & b^{-2} & 0 \\
0 & 0 & c^{-2}
\end{array} \right] \mathbf{V}^{\top} \: ,
\label{eq:A}
\end{equation}
and $\mathbf{V}$ is an orthogonal matrix whose first, second and third columns
are defined by unit vectors $\mathbf{v}_{1}$, $\mathbf{v}_{2}$, and $\mathbf{v}_{3}$
(Fig. \ref{fig:orientation-angles}b), respectively.
These unit vectors are commonly described in terms of auxiliary
orientation angles $\alpha$, $\gamma$, and $delta$, which are not used
by the geoscientific community. These angles, however, can be defined from
$\varepsilon$, $\zeta$, and $\eta$ (Fig. \ref{fig:orientation-angles}a)
as follows \citep{clark1986}:
\begin{equation}
\alpha = \varepsilon - \cos^{-1} \left[ \frac{\cos \eta}
{\left( 1 - \sin^{2} \zeta \, \sin^{2} \eta \right)^{\frac{1}{2}}}\right] \: ,
\label{eq:alpha}
\end{equation}
\begin{equation}
\gamma = \tan^{-1} \left( \frac{\cos \zeta}{\sin \zeta \, \cos \eta}\right)
\label{eq:gamma}
\end{equation}
and
\begin{equation}
\delta = \sin^{-1} \left( \sin \zeta \, \sin \eta \right) \: ,
\label{eq:delta}
\end{equation}
where $-90^{\circ} \leq \gamma \leq 90^{\circ}$ and
$0 \leq \delta \leq 90^{\circ}$.
Then, given the orientation angles $\varepsilon$, $\zeta$, and $\eta$
(Fig. \ref{fig:orientation-angles}a), we define the auxiliary angles
$\alpha$, $\gamma$ and $\delta$ (Eqs. \ref{eq:alpha}, \ref{eq:gamma}, and
\ref{eq:delta}) and, finally, the unit vectors
$\mathbf{v}_{1}$, $\mathbf{v}_{2}$, and $\mathbf{v}_{3}$
(Fig. \ref{fig:orientation-angles}b) according to the ellipsoid type.
For triaxial ellipsoids (i.e., $a > b > c$) and prolate ellipsoids
(i.e., $a > b = c$), we opted for defining these unit vectors
by following \citet{clark1986}:
\begin{equation}
\mathbf{v}_{1} = \left[\begin{array}{c}
                  -\cos\alpha \; \cos\delta \\
                  -\sin\alpha \; \cos\delta \\
                  -\sin\delta
                  \end{array} \right] \: ,
\label{eq:v1_triaxial_prolate}
\end{equation}
\begin{equation}
\mathbf{v}_{2} = \left[\begin{array}{c}
                  \cos\alpha \; \cos\gamma \; \sin\delta + \sin\alpha \; \sin\gamma \\
                  \sin\alpha \; \cos\gamma \; \sin\delta - \cos\alpha \; \sin\gamma \\
                  -\cos\gamma \; \cos\delta
                  \end{array} \right] \: ,
\label{eq:v2_triaxial_prolate}
\end{equation}
\begin{equation}
\mathbf{v}_{3} = \left[\begin{array}{c}
                  \sin\alpha \; \cos\gamma - \cos\alpha \; \sin\gamma \; \sin\delta \\
                  -\cos\alpha \; \cos\gamma - \sin\alpha \; \sin\gamma \; \sin\delta \\
                  \sin\gamma \; \cos\delta
                  \end{array} \right] \: .
\label{eq:v3_triaxial_prolate}
\end{equation}
\citet{emerson1985} calculated the unit vectors $\mathbf{v}_{1}$, $\mathbf{v}_{2}$,
and $\mathbf{v}_{3}$ similarly by using Eqs. \ref{eq:v1_triaxial_prolate},
\ref{eq:v2_triaxial_prolate}, and \ref{eq:v3_triaxial_prolate}, but
with $\gamma = 0^{\circ}$.
Finally, we define the unit vectors $\mathbf{v}_{1}$, $\mathbf{v}_{2}$,
and $\mathbf{v}_{3}$ for oblate ellipsoids (i.e., $a < b = c$) as follows:
\begin{equation}
\mathbf{v}_{1} = \left[\begin{array}{c}
                  -\cos\alpha \; \sin\gamma \; \sin\delta + \sin\alpha \; \cos\gamma \\
                  -\sin\alpha \; \sin\gamma \; \sin\delta - \cos\alpha \; \cos\gamma \\
                  \sin\gamma \; \cos\delta
                  \end{array} \right] \: ,
\label{eq:v1_oblate}
\end{equation}
\begin{equation}
\mathbf{v}_{2} = \left[\begin{array}{c}
                  -\cos\alpha \; \cos\delta \\
                  -\sin\alpha \; \cos\delta \\
                  -\sin\delta
                  \end{array} \right] \: ,
\label{eq:v2_oblate}
\end{equation}
\begin{equation}
\mathbf{v}_{3} = \left[\begin{array}{c}
                  \sin\alpha \; \sin\gamma + \cos\alpha \; \cos\gamma \; \sin\delta \\
                  -\cos\alpha \; \sin\gamma + \sin\alpha \; \cos\gamma \; \sin\delta \\
                  -\cos\gamma \; \cos\delta
                  \end{array} \right] \: .
\label{eq:v3_oblate}
\end{equation}
This approach is very similar to that presented by \citet{emerson1985}.

@@

Based on \citeauthor{maxwell1873}'s postulate, let us
assume that the body has a uniform magnetization given by
\begin{equation}
\mathbf{M} = \mathbf{K} \, \mathbf{H}^{\dagger} \: ,
\label{eq:M-KH}
\end{equation}
where $\mathbf{H}^{\dagger}$ is the resultant uniform magnetic field
at any point within the body and $\mathbf{K}$ is a
constant and symmetrical 2nd-order tensor representing the
magnetic susceptibility of the body.
This is a good approximation for bodies at room temperature,
subjected to an inducing field $\mathbf{H}_{0}$ with
strength $\leq 10^{-3} \mu_{0}^{-1}$ \unit{A \, m^{-1}} \citep{rochette1992},
where $\mu_{0}$ represents the magnetic constant (in \unit{H \, m^{-1}}).
In this case, the susceptibility tensor $\mathbf{K}$ is commonly
represented, in the main coordinate system (Fig. \ref{fig:orientation-angles}a),
as follows:
\begin{equation}
\mathbf{K} = \mathbf{U}
\left[ \begin{array}{ccc}
k_{1} & 0 & 0 \\
0 & k_{2} & 0 \\
0 & 0 & k_{3}
\end{array} \right] \mathbf{U}^{\top} \: ,
\label{eq:K}
\end{equation}
where $k_{1} > k_{2} > k_{3}$ are the
\textit{principal susceptibilities} and $\mathbf{U}$ is
an orthogonal matrix whose columns $\mathbf{u}_{i}$,
$i = 1, 2, 3$, are unit vectors called
\textit{principal directions}.
The unit vectors $\mathbf{u}_{i}$ may be defined
similarly to $\mathbf{v}_{1}$, $\mathbf{v}_{2}$,
and $\mathbf{v}_{3}$ (Eqs. \ref{eq:v1_triaxial_prolate},
\ref{eq:v2_triaxial_prolate}, \ref{eq:v3_triaxial_prolate},
\ref{eq:v1_oblate}, \ref{eq:v2_oblate}, and \ref{eq:v3_oblate}),
by a given set of orientation angles $\varepsilon$, $\zeta$,
and $\eta$ (Eqs. \ref{eq:alpha}, \ref{eq:gamma}, and
\ref{eq:delta}).

@@

%The angular difference $\theta$ between the resultant magnetization
%and the inducing field is the same, whether the quantities
%are defined in the main coordinate system (Fig. \ref{fig:orientation-angles}a)
%or local coordinate system (Fig. \ref{fig:orientation-angles}b).
%For convenience, we define $\theta$ (in degrees) in the main coordinate system
%by using the scalar product definition as follows:
%\begin{equation}
%\theta = \frac{180}{\pi} \, \cos^{-1} \left(
%\frac{\mathbf{M}^{\top}\mathbf{H}_{0}}
%{\| \mathbf{M} \| \; \| \mathbf{H}_{0} \|}\right) \: ,
%\label{eq:theta}
%\end{equation}
%where $\mathbf{M}$ is obtained by pre-multiplying the
%magnetization $\tilde{\mathbf{M}}$ (Eq. \ref{eq:M-tilde-remanence})
%by the orthogonal matrix $\mathbf{V}$ (Eq. \ref{eq:A}),
%which is defined according to the ellipsoid type, and
%$\mathbf{H}_{0}$ is given by Eq. \ref{eq:H0}.
%
For the particular case in which the susceptibility is isotropic,
the susceptibility tensor is defined according to Eq. \ref{eq:K-isotropic}.
In this case, the magnetization $\mathbf{M}$ (Eqs. \ref{eq:M-KH} and \ref{eq:M-remanence}),
referred to the main coordinate system (Fig. \ref{fig:orientation-angles}a),
and the matrix $\mathbf{\Lambda}$ (Eq. \ref{eq:Lambda})
can be rewritten as follows:
\begin{equation}
\mathbf{M} = \mathbf{\Lambda} \left( \chi \, \mathbf{H}_{0} +
\mathbf{M}_{R} \right) \: ,
\label{eq:M-K-isotropic}
\end{equation}
and
\begin{equation}
\mathbf{\Lambda} = \mathbf{V}
\left( \mathbf{I} + \chi \, \tilde{\mathbf{N}}^{\dagger} \right)^{-1}
\mathbf{V}^{\top} \: .
\label{eq:Lambda-K-isotropic}
\end{equation}
Despite the coordinate transformation represented by
matrix $\mathbf{V}$ (Eq. \ref{eq:A}),
this equation is in perfect agreement with those presented
by \citet[Eqs. ~13--15]{guo2001}.
The first term, depending on the inducing field $\mathbf{H}_{0}$
(Eq. \ref{eq:H0}),
represents the induced magnetization whereas the term depending on
$\mathbf{M}_{R}$ is the remanent magnetization.
Equation \ref{eq:M-K-isotropic} reveals that, as pointed out by
many authors
\citep[e.g.,][]{maxwell1873, dubois1896, stoner1945, clark1986, sttraton2007},
the induced magnetization opposes the inducing field
if it is parallel to an ellipsoid axis,
independently of the ellipsoid type. Otherwise, the
magnetization is not necessarily parallel to
the inducing field.
If we additionally consider that $\chi << 1$, the matrix
$\mathbf{\Lambda}$ (Eq. \ref{eq:Lambda-K-isotropic})
approaches to the identity and the magnetization
$\mathbf{M}$ (Eq. \ref{eq:M-K-isotropic})
can be approximated by:
\begin{equation}
\breve{\mathbf{M}} = \chi \, \mathbf{H}_{0} +
\mathbf{M}_{R} \: ,
\label{eq:M-approx}
\end{equation}
which is the classical equation describing the resultant
magnetization in applied geophysics \citep[p. ~89]{blakely1996}.
Notice that, in this particular case, the induced magnetization
is parallel to the inducing field $\mathbf{H}_{0}$ (Eq. \ref{eq:H0}),
whether it is parallel to an ellipsoid axis or not.
Usually, Eq. \ref{eq:M-approx}
is considered a good approximation for $\chi \leq 0.1$ \unit{SI}.
Although this value has been largely used in the literature,
there have been few empirical and/or theoretical investigations
about it.

@@

@@

@@

\citet{farrar1979} represented the Warrego orebody as a triaxial
ellipsoid having a high isotropic susceptibility.
In this case, the self-demagnetization strongly impacts the
magnetic modelling of this body.
According to \citet{wedekind1990}, the Warrego orebody is a combination of
two major and several small ironstone lodes, which are discrete bodies
comprised predominantly of magnetite or hematite above the base of
oxidation.
The Warrego orebody was the resource on which the Warrego mine was
developed. After nearly a decade as one of the most important
gold and copper mines in Australia, the Warrego mine was closed
in late 1989.

Table \ref{tab:parameters-warrego} shows the parameters defining
a synthetic orebody which is based on that presented by
\citet{farrar1979} to represent the Warrego orebody.
Figure \ref{fig:tfa-warrego} shows the total-field anomaly
$\Delta T (\mathbf{r})$ (Eq. \ref{eq:delta-T-approx}) produced
by the synthetic body on a regular grid of $100 \times 100$
points at a constant vertical coordinate $z = 0$ \unit{m}.
The total-field anomaly varies from $\approx -176$ \unit{nT}
to $\approx 237$ \unit{nT}, resulting in a peak-to-peak amplitude
of $\approx 413$ \unit{nT}, and was calculated by using the
true magnetization $\mathbf{M}$ defined in Eqs.
\ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda}.

@@

As expected, the differences calculated by using the higher isotropic
susceptibility (Fig. \ref{fig:tfa-differences-warrego}a) are very large.
The peak-to-peak amplitude is $\approx 141$ \unit{nT}
and represents $\approx 34 \%$ of the peak-to-peak amplitude of
the total-field anomaly shown in Fig. \ref{fig:tfa-warrego}.
This result reinforces that the use of the approximated magnetization
$\breve{\mathbf{M}}$ (Eq. \ref{eq:M-approx}) may negatively impact
the magnetic modelling of bodies with high isotropic susceptibility.

Figure \ref{fig:tfa-differences-warrego}b shows the differences calculated
by using $\chi_{1} = 0.1$ \unit{SI}.
It is commonly accepted that, for bodies having isotropic susceptibilities
lower than or equal to $0.1$ \unit{SI}, the self-demagnetization can be neglected and,
consequently, the magnetization $\breve{\mathbf{M}}$ (Eq. \ref{eq:M-approx})
is a good approximation of the true magnetization $\mathbf{M}$
(Eqs. \ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda}).
In our test, the use of $\chi_{1} = 0.1$ \unit{SI} leads to a relative error
$\| \delta \mathbf{M} \| \, \| \mathbf{M} \|^{-1} \approx 2.4 \%$
(Eq. \ref{eq:perturbed-M-inequality}) in the magnetization.
The peak-to-peak amplitude of the differences in the total-field anomaly
(Fig. \ref{fig:tfa-differences-warrego}b) is $\approx 0.8$ \unit{nT},
which represents $\approx 2.9 \%$ of the peak-to-peak amplitude of the
total-field anomaly calculates by using the true magnetization $\mathbf{M}$
(Eqs. \ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda}).

Finally, Fig. \ref{fig:tfa-differences-warrego}c shows the differences
calculated by using $\chi_{2} = 0.116$ \unit{SI}.
This value was calculated by using Eq. \ref{eq:chi-max} with a
target relative error $\epsilon = 8 \%$ and the $\tilde{n}^{\dagger}_{max}$
defined by Eq. \ref{eq:n-tilde-dagger-33-triaxial}.
By using this isotropic susceptibility, it is expected that the
calculated relative error
$\| \delta \mathbf{M} \| \, \| \mathbf{M} \|^{-1}$
(Eq. \ref{eq:perturbed-M-inequality}) in the magnetization be
lower than or equal to the target relative error $\epsilon = 8 \%$.
In this test, the use of $\chi_{2} = 0.116$ \unit{SI} leads to a relative error
$\| \delta \mathbf{M} \| \, \| \mathbf{M} \|^{-1} \approx 2.8 \%$
(Eq. \ref{eq:perturbed-M-inequality}) in the magnetization.
The peak-to-peak amplitude of the differences in the total-field anomaly
(Fig. \ref{fig:tfa-differences-warrego}c) is $\approx 1.1$ \unit{nT},
which represents $\approx 3.3 \%$ of the peak-to-peak amplitude of the
total-field anomaly calculates by using the true magnetization $\mathbf{M}$
(Eqs. \ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda}).
In this case, the use an isotropic susceptibility greater than the usual
limit $0.1$ \unit{SI} does not mislead the magnetic modelling
dramatically. On the contrary, it shows small discrepancies in the
magnetic modelling and validates Eq. \ref{eq:chi-max}.

>>>>

Let $(x, y, z)$ be a point referred to a Cartesian coordinate system with axes
$x$, $y$ and $z$ pointing to, respectively, North, East and down.
For convenience, we denominate this coordinate system as
\textit{main coordinate system} (Fig. \ref{fig:orientation-angles}a).
Let us consider an ellipsoidal body with centre at the point
$(x_{c}, y_{c}, z_{c})$, orientation defined by the angles
\textit{strike} $\varepsilon$, \textit{dip} $\zeta$, and \textit{rake} $\eta$
(Fig. \ref{fig:orientation-angles}a), and semi-axes defined by
positive constants $a$, $b$, $c$ (Fig. \ref{fig:orientation-angles}b).
The orientation angles \textit{strike}, \textit{dip}, and \textit{rake}
are commonly used to define the orientation of lines in structural geology
\citep{pollard2005, allmendinger2012}.
The points $(x, y, z)$ located on the surface of this ellipsoidal body
satisfy the following equation:
\begin{equation}
(\mathbf{r} - \mathbf{r}_c)^T \mathbf{A} (\mathbf{r} - \mathbf{r}_c) = 1 \: ,
\label{eq:ellipsoid_surface}
\end{equation}
where $\mathbf{r} = [\begin{array}{ccc} x & y & z \end{array} ]^{\top}$,
$\mathbf{r}_{c} = [\begin{array}{ccc} x_{c} & y_{c} & z_{c} \end{array} ]^{\top}$,
$\mathbf{A}$ is a positive definite matrix given by
\begin{equation}
\mathbf{A} = \mathbf{V}
\left[ \begin{array}{ccc}
a^{-2} & 0 & 0 \\
0 & b^{-2} & 0 \\
0 & 0 & c^{-2}
\end{array} \right] \mathbf{V}^{\top} \: ,
\label{eq:A}
\end{equation}
and $\mathbf{V}$ is an orthogonal matrix whose first, second and third columns
are defined by unit vectors $\mathbf{v}_{1}$, $\mathbf{v}_{2}$, and $\mathbf{v}_{3}$
(Fig. \ref{fig:orientation-angles}b), respectively.
The matrix $\mathbf{V}$ can be defined in terms of three rotation matrices:
\begin{equation}
\mathbf{R}_{1}(\theta) = \left[ \begin{array}{ccc}
1 & 0 & 0 \\
0 & \cos \theta & \sin \theta \\
0 & -\sin \theta & \cos \theta
\end{array} \right] \: ,
\label{eq:R1}
\end{equation}
\begin{equation}
\mathbf{R}_{2}(\theta) = \left[ \begin{array}{ccc}
\cos \theta & 0 & -\sin \theta \\
0 & 1 & 0 \\
\sin \theta & 0 & \cos \theta
\end{array} \right]
\label{eq:R2}
\end{equation}
and
\begin{equation}
\mathbf{R}_{3}(\theta) = \left[ \begin{array}{ccc}
\cos \theta & \sin \theta & 0 \\
-\sin \theta & \cos \theta & 0 \\
0 & 0 & 1
\end{array} \right] \: .
\label{eq:R3}
\end{equation}
For triaxial ellipsoids (i.e., $a > b > c$) and prolate ellipsoids
(i.e., $a > b = c$), we define the orthogonal matrix $\mathbf{V}$
as follows:
\begin{equation}
\mathbf{V} =
\mathbf{R}_{1} \left( \frac{\pi}{2} \right) \;
\mathbf{R}_{2} \left( \varepsilon \right) \;
\mathbf{R}_{1} \left( \frac{\pi}{2} - \zeta \right) \;
\mathbf{R}_{3} \left( \eta \right) \: .
\label{eq:V_triaxial_prolate}
\end{equation}
For oblate ellipsoids (i.e., $a < b = c$), we define $\mathbf{V}$ as follows:
\begin{equation}
\mathbf{V} =
\mathbf{R}_{3} \left( -\frac{\pi}{2} \right) \;
\mathbf{R}_{1} \left( \pi \right) \,
\mathbf{R}_{3} \left( \varepsilon \right) \,
\mathbf{R}_{2} \left( \frac{\pi}{2} - \zeta \right) \,
\mathbf{R}_{1} \left( \eta \right) \: .
\label{eq:V_oblate}
\end{equation}
The orthogonal matrices $\mathbf{V}$ used here for triaxial, prolate and oblate
ellipsoids (Eqs. \ref{eq:V_triaxial_prolate} and \ref{eq:V_oblate}) are
different from that used by \citet{emerson1985} and \citet{clark1986}.

@@

Based on \citeauthor{maxwell1873}'s postulate, let us
assume that the body has a uniform magnetization given by
\begin{equation}
\mathbf{M} = \mathbf{K} \, \mathbf{H}^{\dagger} \: ,
\label{eq:M-KH}
\end{equation}
where $\mathbf{H}^{\dagger}$ is the resultant uniform magnetic field
at any point within the body and $\mathbf{K}$ is a
constant and symmetrical 2nd-order tensor representing the
magnetic susceptibility of the body.
This is a good approximation for bodies at room temperature,
subjected to an inducing field $\mathbf{H}_{0}$ with
strength $\leq 10^{-3} \mu_{0}^{-1}$ \unit{A \, m^{-1}} \citep{rochette1992},
where $\mu_{0}$ represents the magnetic constant (in \unit{H \, m^{-1}}).
In this case, the susceptibility tensor $\mathbf{K}$ is commonly
represented, in the main coordinate system (Fig. \ref{fig:orientation-angles}a),
as follows:
\begin{equation}
\mathbf{K} = \mathbf{U}
\left[ \begin{array}{ccc}
k_{1} & 0 & 0 \\
0 & k_{2} & 0 \\
0 & 0 & k_{3}
\end{array} \right] \mathbf{U}^{\top} \: ,
\label{eq:K}
\end{equation}
where $k_{1} > k_{2} > k_{3}$ are the
\textit{principal susceptibilities} and $\mathbf{U}$ is
an orthogonal matrix whose columns $\mathbf{u}_{i}$,
$i = 1, 2, 3$, are unit vectors called
\textit{principal directions}.
Similarly to the matrix $\mathbf{V}$ (Eqs. \ref{eq:ellipsoid_surface},
\ref{eq:V_triaxial_prolate} and \ref{eq:V_oblate}), we define
the matrix $\mathbf{U}$ as a function
of given orientation angles $\varepsilon$, $\zeta$,
and $\eta$ depending on the ellipsoid type. For triaxial
and prolate ellipsoids, we define $\mathbf{U}$ by
using Eq. \ref{eq:V_triaxial_prolate}, whereas for oblate ellipsoids
we use Eq. \ref{eq:V_oblate}.
Notice that the orientation angles $\varepsilon$, $\zeta$,
and $\eta$ defining the orthogonal matrix $\mathbf{U}$
may be different from that angles $\varepsilon$, $\zeta$,
and $\eta$ defining the ellipsoid orientation
(Fig. \ref{fig:orientation-angles}).

@@

For the particular case in which the susceptibility is isotropic,
the susceptibility tensor is defined according to Eq. \ref{eq:K-isotropic}.
In this case, the magnetization $\mathbf{M}$ (Eqs. \ref{eq:M-KH} and \ref{eq:M-remanence}),
referred to the main coordinate system (Fig. \ref{fig:orientation-angles}a),
and the matrix $\mathbf{\Lambda}$ (Eq. \ref{eq:Lambda})
can be rewritten as follows:
\begin{equation}
\mathbf{M} = \mathbf{\Lambda} \left( \chi \, \mathbf{H}_{0} +
\mathbf{M}_{R} \right) \: ,
\label{eq:M-K-isotropic}
\end{equation}
and
\begin{equation}
\mathbf{\Lambda} = \mathbf{V}
\left( \mathbf{I} + \chi \, \tilde{\mathbf{N}}^{\dagger} \right)^{-1}
\mathbf{V}^{\top} \: .
\label{eq:Lambda-K-isotropic}
\end{equation}
Despite the coordinate transformation represented by
matrix $\mathbf{V}$ (Eq. \ref{eq:A}),
this equation is in perfect agreement with those presented
by \citet[Eqs. ~13--15]{guo2001}.
The first term, depending on the inducing field $\mathbf{H}_{0}$
(Eq. \ref{eq:H0}),
represents the induced magnetization whereas the term depending on
$\mathbf{M}_{R}$ is the remanent magnetization.
Equation \ref{eq:M-K-isotropic} reveals that, as pointed out by
many authors
\citep[e.g.,][]{maxwell1873, dubois1896, stoner1945, clark1986, sttraton2007},
the induced magnetization opposes the inducing field
if it is parallel to an ellipsoid axis,
independently of the ellipsoid type. Otherwise, the
magnetization is not necessarily parallel to
the inducing field.
If we additionally consider that $\chi << 1$, the matrix
$\mathbf{\Lambda}$ (Eq. \ref{eq:Lambda-K-isotropic})
approaches to the identity and the magnetization
$\mathbf{M}$ (Eq. \ref{eq:M-K-isotropic})
can be approximated by:
\begin{equation}
\breve{\mathbf{M}} = \chi \, \mathbf{H}_{0} +
\mathbf{M}_{R} \: ,
\label{eq:M-approx}
\end{equation}
which is the classical equation describing the resultant
magnetization in applied geophysics \citep[p. ~89]{blakely1996}.
Notice that, in this particular case, the induced magnetization
is parallel to the inducing field $\mathbf{H}_{0}$ (Eq. \ref{eq:H0}),
whether it is parallel to an ellipsoid axis or not.
Usually, Eq. \ref{eq:M-approx}
is considered a good approximation for $\chi \leq 0.1$ \unit{SI}.
Although this value has been largely used in the literature,
there have been few empirical and/or theoretical investigations
about it.

@@

All the code developed for generating the results presented in the following
subsections, as well as the code developed for generating additional numerical simulations,
can be found at the folder \textit{code} of the repository
\url{https://github.com/pinga-lab/magnetic-ellipsoid}.

@@

The code used for generating the results presented in this section
can found at \url{https://github.com/pinga-lab/magnetic-ellipsoid/code}

@@

We simulated an ellipsoidal body similar to the Warrego orebody,
which was the resource on which the well-known Warrego mine was
developed in Tennant Creek, Australia. After nearly a decade as
one of the most important gold and copper mines in Australia, the
Warrego mine was closed in late 1989.
According to \citet{wedekind1990}, the Warrego orebody is a combination of
two major and several small ironstone lodes, which are discrete bodies
comprised predominantly of magnetite or hematite above the base of
oxidation.
\citet{farrar1979} represented the Warrego orebody as a triaxial
ellipsoid having a high isotropic susceptibility.
In this case, the self-demagnetization strongly impacts the
magnetic modelling of this body.

Table \ref{tab:parameters-warrego} shows the parameters defining
a synthetic orebody which is based on that presented by
\citet{farrar1979} to represent the Warrego orebody.
Figure \ref{fig:tfa-warrego} shows the total-field anomaly
$\Delta T (\mathbf{r})$ (Eq. \ref{eq:delta-T-approx}) produced
by the synthetic body on a regular grid of $100 \times 100$
points at a constant vertical coordinate $z = 0$ \unit{m}.
The total-field anomaly varies from $\approx -71$ \unit{nT}
to $\approx 482$ \unit{nT}, resulting in a peak-to-peak amplitude
of $\approx 553$ \unit{nT}, and was calculated by using the
true magnetization $\mathbf{M}$ defined in Eqs.
\ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda}.

@@

As expected, the differences calculated by using the higher isotropic
susceptibility (Fig. \ref{fig:tfa-differences-warrego}a) are very large.
The peak-to-peak amplitude is $\approx 40$ \unit{nT}
and represents $\approx 8 \%$ of the peak-to-peak amplitude of
the total-field anomaly shown in Fig. \ref{fig:tfa-warrego}.
%This result reinforces that the use of the approximated magnetization
%$\breve{\mathbf{M}}$ (Eq. \ref{eq:M-approx}) may negatively impact
%the magnetic modelling of bodies with high isotropic susceptibility.

Figure \ref{fig:tfa-differences-warrego}b shows the differences calculated
by using $\chi_{1} = 0.1$ \unit{SI}.
It is commonly accepted that, for bodies having isotropic susceptibilities
lower than or equal to $0.1$ \unit{SI}, the self-demagnetization can be neglected and,
consequently, the magnetization $\breve{\mathbf{M}}$ (Eq. \ref{eq:M-approx})
is a good approximation of the true magnetization $\mathbf{M}$
(Eqs. \ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda}).
In our test, the use of $\chi_{1} = 0.1$ \unit{SI} leads to a relative error
$\| \delta \mathbf{M} \| \, \| \mathbf{M} \|^{-1} \approx 0.7 \%$
(Eq. \ref{eq:perturbed-M-inequality}) in the magnetization.
The peak-to-peak amplitude of the differences in the total-field anomaly
(Fig. \ref{fig:tfa-differences-warrego}b) is $\approx 0.2$ \unit{nT},
which represents $\approx 0.6 \%$ of the peak-to-peak amplitude of the
total-field anomaly calculates by using the true magnetization $\mathbf{M}$
(Eqs. \ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda}).

Finally, Fig. \ref{fig:tfa-differences-warrego}c shows the differences
calculated by using $\chi_{2} = 0.116$ \unit{SI}.
This value was calculated by using Eq. \ref{eq:chi-max} with a
target relative error $\epsilon = 8 \%$ and the $\tilde{n}^{\dagger}_{max}$
defined by Eq. \ref{eq:n-tilde-dagger-33-triaxial}.
By using this isotropic susceptibility, it is expected that the
calculated relative error
$\| \delta \mathbf{M} \| \, \| \mathbf{M} \|^{-1}$
(Eq. \ref{eq:perturbed-M-inequality}) in the magnetization be
lower than or equal to the target relative error $\epsilon = 8 \%$.
In this test, the use of $\chi_{2} = 0.116$ \unit{SI} leads to a relative error
$\| \delta \mathbf{M} \| \, \| \mathbf{M} \|^{-1} \approx 0.8 \%$
(Eq. \ref{eq:perturbed-M-inequality}) in the magnetization.
The peak-to-peak amplitude of the differences in the total-field anomaly
(Fig. \ref{fig:tfa-differences-warrego}c) is $\approx 0.3$ \unit{nT},
which represents $\approx 0.7 \%$ of the peak-to-peak amplitude of the
total-field anomaly calculates by using the true magnetization $\mathbf{M}$
(Eqs. \ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda}).
In this case, the use an isotropic susceptibility greater than the usual
limit $0.1$ \unit{SI} does not mislead the magnetic modelling
dramatically. On the contrary, it shows small discrepancies in the
magnetic modelling and validates Eq. \ref{eq:chi-max}.
